I'm going to be taking a look at the New York Knicks players advanced statistics for the 2022/2023 season.

```{r libraries}

library(tidyverse)
library(rvest)
library(dplyr)
library(ggplot2)
library(XML)
library(ggthemes)

```

Firstly, we're going to have to webscrape. We'll be looking at Knicks counting stats first.

```{r webscrape}
myLink1 <- "https://www.basketball-reference.com/teams/NYK/2023.html"
page1 <- read_html(myLink1)
tables <- html_table(page1)

counting_stats <- tables[[2]]
head(counting_stats)

#Now it's webscraped! Hooray! Our next step, is to make this data look nice.

counting_stats <- counting_stats %>% 
  select(-Rk)

#I want to add position to each player as well, so I'll need to webscrape the first page off this table.

positions <- tables[[1]] %>% 
  select(Player, Pos)

#Let's make the data sets more compatible for merging

positions[14, "Player"] <- "Trevor Keels"
positions[15, "Player"] <- "Duane Washington Jr."

counting_stats <- merge(counting_stats, positions, by = "Player", all = TRUE)

#Let's adjust it again and remove players with no games played and add positions by our own Google search for those who need it

counting_stats <- counting_stats[!is.na(counting_stats$G), ] %>% 
  select(Player, Pos, Age, everything())

counting_stats[1, "Pos"] <- "SF"
counting_stats[15, "Pos"] <- "PG"
counting_stats[16, "Pos"] <- "SF"
```

Now, let's do the same thing but for the advanced statistics.

```{r advanced stats}

advanced_stats <- tables[[6]]

#Let's clean it up a little bit.

advanced_stats <- advanced_stats %>% 
  select(-c(18, 23))

```

I'm going to create a violin plot to compare the impact of the starters in comparison to players coming off the bench. I'm going to use Box Plus/Minus in the advanced statistic data frame. This statistic is adjusted per 100 possessions, so there is less of a bias for the starters (because they play more)

```{r ggplot, warning = FALSE}

#Let's first determine the starters versus the bench. I'm going to create a variable called GS_P - Games Started Percentage

counting_stats$GS_P <- counting_stats$GS / counting_stats$G

start_bench <- counting_stats %>% 
  select(Player, Pos, Age, G, GS, GS_P, MP)

#I'm going to exclude players based on several different categories. Firstly, games played. I want players who got minutes for a majority of the season or are Josh Hart. Josh Hart was a trade deadline acquisition for the Knicks and became an invaluable piece of the team. The arbitrary number I've decided upon is 20 games.

start_bench <- start_bench %>%
  filter(!(G < 20 & Player != "Josh Hart")) #ChatGPT consulted

#Next, we're going to make a category for whether we consider the player a starter or a bench player. I'm going to use the percentage of 80%, leaving us with five starters and six bench players.

start_bench <- start_bench %>% 
  mutate(Role = ifelse(GS_P >= 0.8, "Starter", "Bench"))

#Merging the statistic of choice from the advanced data set

start_bench <- merge(start_bench, advanced_stats, by = "Player")  

start_bench <- start_bench %>% 
  select(Player, Age.x, G.x, GS, GS_P, MP.x, Role, BPM) %>% 
  rename(Age = Age.x) %>% 
  rename(G = G.x) %>% 
  rename(MP = MP.x)

#Now, we can begin plotting

ggplot(start_bench, aes(x = Role, y = BPM, fill = Role)) +
  geom_violin() +
  geom_boxplot(width = 0.2, fill = "white", color = "black", outlier.shape = NA) + 
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  labs(title = "BPM of Starters vs Bench, Knicks 22/23", x = "Role", y = "BPM") +
  theme_minimal()



```

Of course, we wish the sample size was higher but only so many players can play for a basketball season. 

This violin plot, shows the difference in quality between the starters and the bench. The media billed the Knicks' bench as one of the best in the league and there are various metric that back it. However, the starters are still superior. This Box Plus Minus statistic compares each Knicks athlete to the league average. Hopefully, they would be above and they are for the most part (besides RJ Barrett unsurprisingly - we have a love-hate relationship).

```{r ggplot scatter, warning = FALSE}

#We're going to make a scatter plot comparing the Defensive Box Plus Minus and their PER, player effeciency rating, for the starters and bench players we looked at in the violin plot. A PER of 15 is a league-average player.

advanced_stats_merge <- advanced_stats %>% 
  select(Player, DBPM, PER)

per_knicks <- merge(start_bench,advanced_stats_merge, by = "Player")

per_knicks$Last_Name <- sapply(strsplit(per_knicks$Player, " "), function(x) x[2]) #ChatGPT consulted

ggplot(per_knicks, aes(PER, DBPM, color = Role)) +
  geom_point() +
  geom_text(aes(label = Last_Name), vjust = -0.5) +
  labs(title = "Knicks Starters 2022/23",
       x = "PER",
       y = "Defensive Box Plus Minus") +
  scale_x_continuous(limits = c(4.8, 23)) + 
  scale_y_continuous(limits= c(-2, 2.5)) +
  geom_vline(xintercept = 15, color = "red", linetype = "dashed") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  annotate("text", x = 0, y = -2.5, label = "PER of 15 and DBPM of 0 are League Average", hjust = 0, vjust = 0) +
  theme_fivethirtyeight() +
  theme(axis.title.x = element_text(margin = margin(t = 7)),
        axis.title.y = element_text(margin = margin(r = 8))) +
  theme(plot.caption = element_text(hjust = 0, size = 8, margin = margin(t = 15))) +
  labs(caption = "PER of 15 and DBPM of 0 are League Average")
  
```
In this graphic, we can compare the defensive abilities of all Knicks starters and their PER. The PER statistic does take counting statistics into account so its biased for the starters. However, DBPM does not and you see the true strength of the Knicks bench. they are very strong on defense. Excluding Rose, who only played 34 games this season, the only negative defensive presence on the Knicks bench this season was Obi Toppin. Toppin was nearly average with PER while only playing 15.7 minutes per game - his offensive contributions made up for his lack of defense. The violin plot above misses the true strength of the Knicks bench: they're defensive capabilites.

